---
title: "Yelp-Dataset-Project"
author: "Ana Perez, Clarissa Vazquez, Naelin Aquino"
date: "May 10, 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(prompt=TRUE, comment="", echo=TRUE)
```

```{r}
#preparing the datasets
source("process-data.r")

fnames = c('business','checkin','review','tip','user')
jfile = paste0('~/yelp_dataset_challenge_round9/yelp_academic_dataset_',fnames,'.json')

business_dat = business_data(jfile[1])
checkin_dat = checkin_data(jfile[2])
review_dat = review_data(jfile[3])
tip_dat = tip_data(jfile[4])
user_dat = user_data(jfile[5])


#scale user data

#scale business data

#scale checkin data

#scale tip data
```

## User Data Exploration

```{r}
hist(user_dat$elite[user_dat$elite != 0], xlab = "years", ylab = "people", main="Elite Yelpers", col = "purple", xlim = c(2, 13))
```

Pictured in this histogram is the amount of years a Yelp user has been an Elite Yelper. An Elite Yelper is someone who gives well-written reviews, high quality tips, have a detailed personal profile, an active voting and complimenting record, and a history of playing well with others. Out of 1029432 about 44246 yelpers are Elite. 

## Business Data Exploration

```{r}
barplot(table(business_dat$state), horiz= TRUE, las = 1, xlim = c(0,50000), cex.names = 0.7, space = 1, col = "purple", main = "Amount of Business By State")
```

Most of the business from this dataset come from Arizona, Nevada, and Ontario Canada.

## Checkin Data Exploration

```{r}

```

## Reviews Data Exploration

```{r}

```

## Tip Data Exploration




